Final Project

### 2025-05-03 MSSC5750_Project.qmd
More Visualizations/Transformations

```{r echo = FALSE, message=FALSE}
library(MASS)
library(ggplot2)
library(dplyr)
library(scales)
library(glmnet)
library(fastDummies)
library(caret)
library(shapr)
library(shapviz)
library(party)
library(reshape2)

@@

@@

#Actually might be better to just add a catch bucket at the end. IF over 250k, just set to 250k
ggplot(data=data %>% mutate(person_income2 = if_else(person_income<250000,person_income,250000)), aes(x=person_income2)) +geom_histogram(bins=40, fill='steelblue',color='black') + xlab("Income") + ggtitle("Histogram of Income")+
  scale_x_continuous(breaks=seq(0,250000,50000), labels=c("$0","$50,000","$100,000","$150,000","$200,000","$250,000+") )

temp<-data %>% summarize('0'=1-mean(loan_status),"1"=mean(loan_status) ,.by=loan_intent)
intent<-melt(temp) %>% dplyr::rename(loan_status=variable)

ggplot(data=intent, aes(x=loan_intent, y=value, group=loan_status, fill=loan_status)) + geom_col(position='stack')

>>>>

```{r echo = FALSE, message=FALSE}
library(MASS)
library(ggplot2)
library(dplyr)
library(scales)
library(glmnet)
library(fastDummies)
library(caret)
library(shapr)
library(shapviz)
library(party)
library(reshape2)
library(gridExtra)

@@

# Looking at age/credit score with loan approval. NO clear patterns
ggplot(data=rbind( data %>% filter(loan_status==1) %>%  sample_n(1000), data %>% filter(loan_status==0) %>%  sample_n(1000)), aes(x=credit_score, y=person_age, fill=loan_status, color=loan_status)) + 
  geom_point() + ylim(20,60) + xlim(400,780)

@@

#Actually might be better to just add a catch bucket at the end. IF over 500k, just set to 500k
inc1<-ggplot(data=data %>% mutate(person_income2 = if_else(person_income<500000,person_income,500000)), aes(x=person_income2)) +geom_histogram(bins=40, fill='steelblue',color='black') + xlab("Income") + ggtitle("Histogram of Income") +
  scale_x_continuous(breaks=seq(0,500000,50000), labels=c("$0","$50,000","$100,000","$150,000","$200,000","$250,000","$300,000","$350,000","$400,000","$450,000", "$500,000+") )

# Visual for log adjusted income
inc2<-ggplot(data = data %>% mutate('log(person_income)' = log(ifelse(person_income>500000,500000,person_income))  ), aes(x=log(person_income))) +geom_histogram(bins=40, fill='orange',color='black') + xlab("Log of Income") + ggtitle("Histogram of Log of Income")

grid.arrange(inc1, inc2, nrow = 2)

# Loan intent
temp<-data %>% summarize('0'=1-mean(loan_status),"1"=mean(loan_status) ,.by=loan_intent)
intent<-melt(temp) %>% dplyr::rename(loan_status=variable)

ggplot(data=intent %>% filter(loan_status==1), aes(x=loan_intent, y=value)) + 
  geom_col(position='stack', fill="violetred2") + 
  geom_text(aes(label = percent(value,accuracy=1)), 
            position = position_stack(vjust = 0.5), 
            size = 4) + ggtitle('Loan Repayment by Intent') + 
  scale_y_continuous(name='', labels = scales::percent, limits=c(0,.4)) + xlab('') +
  theme(panel.grid.minor.y=element_blank())

# Education
temp<-data %>% summarize('0'=1-mean(loan_status),"1"=mean(loan_status) ,.by=person_education)
educ<-melt(temp) %>% dplyr::rename(loan_status=variable)

ggplot(data=educ %>% filter(loan_status==1), aes(x=person_education, y=value)) +
    geom_col(position='stack', fill='aquamarine3') +
    geom_text(aes(label = percent(value,accuracy=1)), 
            position = position_stack(vjust = 0.5), 
            size = 4) + ggtitle('Loan Repayment by Education') + 
  scale_y_continuous(name='', labels = scales::percent, limits=c(0,.4)) + xlab('') +
  theme(panel.grid.minor.y=element_blank())

### cred hist length
ggplot(data=data %>% summarize(count=n(),.by=cb_person_cred_hist_length), aes(x= cb_person_cred_hist_length, y=count)) + geom_col(fill='steelblue')
